---
title: "ISA 444: Business Forecasting"
subtitle: "02 - Introduction to Time Series Analysis"
author: Fadel M. Megahed
date: "January 27, 2021"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
    number_sections: TRUE
    theme: spacelab
    paged_df: TRUE
    code_folding: show
    code_download: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      cache = FALSE,
                      progress = FALSE, 
                      verbose = FALSE,
                      dpi = 600,
                      dev = 'png',
                      out.width = '100%')
```

# Two Quick Comments Based on Last Class {.tabset .tabset-fade .tabset-pills}

## Beautifying any data frame output {-}

We can capitalize on the [DT](https://rstudio.github.io/DT/) package to beautify any data frame output in R Markdown. For the purpose of being quick, let us compare printing out the built in `iris` dataset using print and the `DT::datatable()` function. 
```{r sampleOutput}
pacman::p_load(tidyverse, DT) # loading the tidyverse and DT packages

iris %>% head(2) # showing only the first two rows for space purposes

datatable(iris, # name of data frame to be printed 
          extensions = c('FixedColumns', 'Buttons'), # extensions used
          options = list(
            dom = 'Bfrtip', # B is the buttons extension, f filtering input, see https://datatables.net/reference/option/dom
            scrollX = TRUE, # scrolling along the x-axis
            buttons = c('copy', 'csv', 'excel', 'pdf'), # Buttons used from the extensions command in Line 40
            fixedColumns = list(leftColumns = 1)) ) %>% # fixing left 1 column due to FixedColumns (similar to viewPane in Excel)
  formatRound(columns= c('Sepal.Length', 'Sepal.Width'), digits=0) # from DT, specifying formatting to specific cols
```


## Potential Issue for Mac Users with the tidyquant package {-}

After last class, one of your colleagues informed me that she had issues in installing the [timetk](https://business-science.github.io/timetk/) package on her Macbook Pro. Given that I do not have a MacBook, I am not able to replicate the error. However, if you are one of these students you have the following options:  

**Plan A:**

  (1) Ensure that your MacOS is up to date (see [how to update the software on your Mac](https://support.apple.com/en-us/HT201541) for more details).  Note that per the [CRAN Documentation](https://cran.r-project.org/bin/macosx/):    
    (a) R 4.0.3 binary for macOS 10.13 (High Sierra) and higher, signed and notarized package. Contains R 4.0.3 framework, R.app GUI 1.73 in 64-bit for Intel Macs, Tcl/Tk 8.6.6 X11 libraries and Texinfo 6.7. The latter two components are optional and can be ommitted when choosing "custom install", they are only needed if you want to use the tcltk R package or build package documentation from sources.  
    (b) Note: the use of X11 (including tcltk) requires [XQuartz](http://xquartz.macosforge.org/) to be installed since it is no longer part of OS X. Always re-install XQuartz when upgrading your macOS to a new major version.    
    (c) **Important:** this release uses Xcode 10.1 and GNU Fortran 8.2. If you wish to compile R packages from sources, you will need to download and GNU Fortran 8.2 - see the [tools](https://cran.r-project.org/bin/macosx/tools/) directory.  
    
  (2) Once you have updated your MacOS software, I would recommend uninstalling both RStudio and R (in that order).  
  (3) Then, I would recommend reinstalling R and RStudio (in that order).  
  (4) Now, you can attempt to re-install the packages that we used last class (`pacman`, `tidyverse`, `tidyquant`). Hopefully, you will be able to install tidyquant.  
  (5) If not, I would recommend installing an earlier version of timetk per [this GitHub Issue Response](https://github.com/business-science/timetk/issues/70#issuecomment-764727521). This can be done as follows:   
    (a) `remove.packages(timetk)` -- assuming it was installed incorrectly.  
    (b) `install.packages('https://cran.r-project.org/src/contrib/Archive/timetk/timetk_2.6.0.tar.gz', repos=NULL, type="source")`, which may be the only step you need if (a) is not needed/ returns an error. 
    
    
**Plan B:**  
I wrote a function that would mimic the output needed from the `tidyquant::tq_get()`. This function can be brought into your working directory using the following command: `source('https://raw.githubusercontent.com/fmegahed/businessForecasting/master/custom_functions/tq_get.R')`. **Note that you will have to do that for every R session, i.e, you can consider this command to be in lieu of the `pacman::p_load(tidyquant)` command.**


# Visualizing COVID-19 Data